$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-05-03 paper/paper.tex
All the way to subject congruence

Weakening states that adding new variables to the typing and input and output usage contexts preserves the well-typedness of a process.
The usage annotations of the new variables will be left untouched as leftovers.

\begin{nitheorem}
  \label{thm:weakening}
  Let $P$ be a well-typed process $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Insert an arbitrary type $t$ at position $i$ in the typing context, and an arbitrary usage annotation $x$ at position $i$ into both the input and output contexts.
  Increment every referrence to a variable greater than or equal to $i$ in process $P$.
  The resulting process will be well-typed so that $\types{insert_i \; t \; \gamma}{insert_i \; x \; \Gamma}{lift_i \; P}{insert_i \; x \; \Xi}$.
\end{nitheorem}
\begin{proof}
  By induction on \textsc{Process} and \textsc{Var}.
\end{proof}

@@

Strengthening states that removing unused variables from the typing and usage contexts preserves the well-typedness of a process.
The proof that a variable is unused within a process must be syntactical and cannot rely on the preservation of its usage annotations between the input and output contexts: a process can refer to a variable and still make no use of it -- e.g. by sending none of its multiplicities over a channel.

\begin{nitheorem}
  \label{thm:strengthening}
  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Let $i$ be a variable index that is not mentioned by $P$, such that $\Unused_i \; P$.
  Then removing index $i$ from the typing context and both of the usage contexts and decrementing every variable in $P$ greater than $i$ preserves its well typedness, so that $\types{delete_i \; \gamma}{delete_i \; \Gamma}{lower_i \; P}{delete_i \; \Xi}$.
\end{nitheorem}
\begin{proof}
  By induction on \textsc{Process} and \textsc{Var}.
\end{proof}

@@

Swapping two variables (in the typing context, input and output usage contexts and process) preserves the well-typedness of a process.

>>>>

Adding a new variable to the context preserves the well-typedness of a process as long as the usage annotation of that variable is preserved unchanged as a leftover.

\begin{nitheorem}
  \label{thm:weakening}
  Let $P$ be a well-typed process $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Insert an arbitrary type $t$ at position $i$ in the typing context, and an arbitrary usage annotation $x$ at position $i$ into both the input and output contexts.
  For simplicity, we overload $insert$ and use it to insert types in typing contexts and usage annotations in usage contexts.
  Increment every variable referrence greater than or equal to $i$ in process $P$.
  The resulting process is well-typed so that $\types{insert_i \; t \; \gamma}{insert_i \; x \; \Gamma}{lift_i \; P}{insert_i \; x \; \Xi}$.
\end{nitheorem}
\begin{proof}
  By induction on \textsc{Process} and \textsc{Var}.
\end{proof}

@@

Removing an unused variable preserves the well-typedness of a process.
The proof that a variable is unused within a process must be syntactical and cannot rely on the preservation of its usage annotations as leftovers: a process can refer to a variable and still make no use of it --- by sending $\lz$ usage annotations over a channel.

\begin{nitheorem}
  \label{thm:strengthening}
  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Let $i$ be a variable index not mentioned by $P$, such that $\Unused_i \; P$.
  Remove the elements at index $i$ from the typing and usage contexts and decrement every variable in $P$ greater than $i$.
  For simplicity, we overload $delete$ and use it to delete types from typing contexts and usage annotations from usage contexts.
  The result preserves the well-typedness of the process, so that $\types{delete_i \; \gamma}{delete_i \; \Gamma}{lower_i \; P}{delete_i \; \Xi}$.
\end{nitheorem}
\begin{proof}
  By induction on \textsc{Process} and \textsc{Var}.
\end{proof}

@@

Swapping two variables preserves the well-typedness of a process.
